---
format:
    html:
        css: styles.css
---

## Objectives {.unnumbered}

## Task 1: `map`, `map2` and `pmap`

```{r}
library(tidyverse)
library(palmerpenguins)
set.seed(123)
list_of_vectors <- list(
    rnorm(20),
    rnorm(20, mean=1),
    rnorm(20, mean=-1, sd=2)
)
numbers <- 1:10
df <- tibble(a = rnorm(10), b=rnorm(10), c=rnorm(10))
summary_stats <- list(mean, median, sd, var)
names <- c("Alice", "Bob", "Charlie")
greetings <- c("Hello", "Hi", "Howdy")
tibbles <- list(
    df1 = tibble(x=1:3, y=5:7),
    df2 = tibble(a=rnorm(5), b=rnorm(5, mean=1)),
    df3 = penguins
)
tibbles.names = c("numbers", "random", "penguins")
variables <- c("flipper_length_mm", "bill_length_mm", "bill_depth_mm")
```

Using `purrr` `map` and `reduce` functions, solve the following coding tasks.

### `map` functions

- [ ] Calculate the mean of each vector in `list_of_vectors`
- [ ] For each number in `list_of_numbers`, create a logical vector with a 1 where the number is even and a zero where it is false. (*Hint*: remember `ifelse`)
- [ ] For each number `<n>` in `list_of_numbers`, create a string that says "Value: <n>"..
- [ ] For each statistic in `summary_stats`, apply the statistic to `numbers`.

::: {.solution}

```{r}
map_dbl(list_of_vectors, mean)
map_int(numbers, \(x) ifelse(x %% 2 == 0, 1, 0))
map_chr(numbers, \(x) paste("Value:",x))
map_dbl(summary_stats, \(x) x(numbers))

```

:::

### `map2` and `pmap`

- [ ] For each `<name>` in `names` and each `<greeting>` in `greetings`, create a string that says `<greeting> <name>`
- [ ] Write each `tibble` in `tibbles` to a file with the corresponding name in `tibbles.names` in the directory `results/`. (*Hint*: use the function `write_csv`).
- [ ] For each `<variable>` in `variables`, produce a plot of that variable against `body_mass_g` in `penguins`. 
- [ ] Create a list of display names for the $y$-axes and add them to your plot.
- [ ] Create a list of titles and add them to your plot.


::: {.solution}

```{r}
pmap_chr(list(greetings, names), \(greeting, name) paste(greeting, name))
pwalk(list(tibbles, tibbles.names), \(tibble, name) {
          path = paste0("results/", name)
          dir.create(dirname(path), recursive=T, showWarnings= F)
          write_csv(tibble, path)
          }
         )
variables <- c("flipper_length_mm", "bill_length_mm", "bill_depth_mm")

ydisplays <- c("Flipper Lenghth", "Bill Length", "Bill Depth")
titles <- c("Flipper Length vs BM", "Bill length vs BM", "Bill depth vs BM")
plots <- pmap(list(variables, ydisplays, titles), \(y,ydisplay, title ) penguins |> ggplot(aes(x = body_mass_g, y = !!sym(y))) +
     geom_point() +
     labs(
      x = "Body Mass",
      y = ydisplay,
      title = title
  )

)
plots[[1]]
plots[[2]]
plots[[3]]
```

:::

## Task 2: `reduce`

```{r} 
# set up code
functions <- list(
    \(x) x + 1,
    \(x) x * 2,
    \(x) sqrt(x)
)

daily_returns <- rnorm(365, mean = 0.001, sd = 0.01)
```

- [ ] Compute the sum and product of `numbers`.
- [ ] Apply the composition of all functions in the list `functions` to a starting number. *Hint:* use the `init` keyword of `reduce`.
- [ ] The `daily_returns` vector represents one year of percentage changes in portfolio value each day. Assuming all the daily returns are reinvested and we start with £100, calculate the portfolio value after one year.

::: {.solution}

```{r}
reduce(numbers, \(agg, curr) agg + curr)
reduce(numbers, \(agg, curr) agg * curr)
starting <- 17
reduce(functions, 
  \(agg, curr) curr(agg),
.init = starting 
)
st_value <- 100
reduce(daily_returns, \(agg, curr) agg * (1 + curr) ,
  .init = st_value
       )
```

:::


## Task 3: Data Frame Mapping Functions

The next two subtasks require you to build *data frames*.
To accomplish this there are two `purrr` functions you can use:

- `map_dfr`: takes a function that returns *dataframe rows* and binds them into a tibble.
- `map_dfc`: as above, but for columns.

Select the appropriate function to rewrite the below loop-based code into a version that uses `purrr`.

```{r}
library(palmerpenguins)

species <- unique(penguins$species)
results <- list()

for (sp in species) {
    data <- penguins |> filter(species == sp)

    # fit a linear regression model of bill depth vs. bill length
    model <- lm(bill_depth_mm ~ bill_length_mm, data=data)

      # Extract the slope and intercept
    slope <- coef(model)["bill_length_mm"]
    intercept <- coef(model)["(Intercept)"]
    
    # Store the results
    results[[sp]] <- tibble(
        species = sp,
        slope = slope,
        intercept = intercept
    )
}

# put into a data frame
loop_results <- bind_rows(results)
loop_results
```

::: {.solution}

```{r}
species <- unique(penguins$species)
map_dfr(species, 
\(sp) {
    data <- penguins |> filter(species == sp)

    # fit a linear regression model of bill depth vs. bill length
    model <- lm(bill_depth_mm ~ bill_length_mm, data=data)

      # Extract the slope and intercept
    slope <- coef(model)["bill_length_mm"]
    intercept <- coef(model)["(Intercept)"]
    
    # Store the results
    tibble(
        species = sp,
        slope = slope,
        intercept = intercept
    )

  }
)
```

:::
