---
title: "Machine Learning Helper functions"
format:
  html:
    self-contained: true
    toc: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(tidymodels)
library(knitr)
library(glmnet)
library(discrim)
library(naivebayes)
library(rpart)
library(rpart.plot)
library(kernlab)
library(nnet)
library(janitor)
library(fairml)
library(ISLR2)
```

### Step 1 - Split Data


```{r}

# Function to split data into train, test, and optionally validation
make_data_split <- function(data,
                            train = 0.8,
                            val = NULL,
                            test = NULL,
                            strata = NULL) {
  data <- data |> drop_na()


  # Split data into validation if needed
  if (is.null(val)) {
    split <- initial_split(
      data,
      prop = train,
      strata = {{ strata }}
    )

    return(list(
      type = "initial_split",
      train = training(split),
      test = testing(split)
    ))
  }

  # Ensure incorrect splits are not provided when explicitly mentioned
  if (is.null(test)) {
    test <- 1 - train - val
  }

  if (abs(train + val + test - 1) > 1e-8) {
    stop("`train`, `val`, and `test` must sum to 1.")
  }

  # Validation split if asked for 
  split <- initial_validation_split(
    data,
    prop = c(train, val),
    strata = {{ strata }}
  )

  list(
    type = "initial_validation_split",
    validation_resamples = validation_set(split),
    train = training(split),
    validation = validation(split),
    test = testing(split)
  )
}
```
#### Step 1 - Use Example:
```{r}
#splits <- make_data_split(penguins, train = 0.8)
#
#splits$train
#splits$test

```

### Step 2 - Preprocessing - Creating Recipe

#### 2.1 - Normal Recipe
```{r}

# Function to create recipe
make_recipe <- function(training_data,
                        outcome,
                        id_vars = NULL) {
  # Outcome variable
  outcome <- enquo(outcome)

  # building recipe with all predictors included
  rec <- recipe(
    training_data,
    formula = as.formula(paste(quo_name(outcome), "~ ."))
  )

  # remove predictors mentioned in the id_vars parameter
  if (!is.null(id_vars)) {
    rec <- rec |>
      update_role(
        all_of(id_vars),
        new_role = "ID"
      )
  }

  rec |>
    step_normalize(all_numeric_predictors()) |>
    step_dummy(all_nominal_predictors())
}
```
#### 2.2 tree recipe

```{r}
# Function to construct tree recipe. Key is not to normalise and create dummy variables
make_tree_recipe <- function(training_data,
                             outcome,
                             id_vars = NULL) {
  outcome <- enquo(outcome)

  rec <- recipe(
    training_data,
    formula = as.formula(paste(quo_name(outcome), "~ ."))
  )

  if (!is.null(id_vars)) {
    rec <- rec |>
      update_role(
        all_of(id_vars),
        new_role = "ID"
      )
  }

  rec
}
```
#### 2.3 Adding pred recipe

```{r}
# Function to create recipe with provided predictors only
make_recipe_addpred <- function(training_data,
                                outcome,
                                preds = NULL,
                                id_vars = NULL) {
  outcome <- enquo(outcome)

  # Build with the provided predictors only
  rhs <- if (is.null(preds)) {
    "."
  } else {
    paste(preds, collapse = " + ")
  }

  rec <- recipe(
    training_data,
    formula = as.formula(
      paste(quo_name(outcome), "~", rhs)
    )
  )

  # Assign ID roles if supplied
  if (!is.null(id_vars)) {
    rec <- rec |>
      update_role(
        all_of(id_vars),
        new_role = "ID"
      )
  }

  rec |>
    step_normalize(all_numeric_predictors()) |>
    step_dummy(all_nominal_predictors())
}
```
#### Step 2 - Use


```{r}
#penguins_recipe <- make_recipe(
#  training_data = penguins_train,
#  outcome = body_mass_g,
#  id_vars = c("island", "sex", "year")
#)
#
#penguins_recipe
```

### Step 3 - Model Selection

#### 1. `linear_reg` model with `lm` engine
```{r}

# Selecting lm model
make_lm_model <- function() {
  linear_reg() |>
    set_engine("lm")
}
```

#### 2. Tuned lasso - `linear_reg` with `glmnet`

```{r}
# Creating lasso model
make_lasso_model <- function() {
  linear_reg(
    penalty = tune(),   # lambda
    mixture = 1         # LASSO (L1)
  ) |>
    set_engine("glmnet")
}
```

#### 3. Tuned ridge - `linear_reg` with `glmnet`

```{r}
# Creating ridge model
make_ridge_model <- function() {
  linear_reg(
    penalty = tune(),   # lambda
    mixture = 0         # Ridge (L2)
  ) |>
    set_engine("glmnet")
}
```
#### 4. Tuned elasticnet - `linear_reg` with `glmnet`


```{r}
# Creating elasticnet model
make_elasticnet_model <- function() {
  linear_reg(
    penalty = tune(),  # lambda
    mixture = tune()   # alpha
  ) |>
    set_engine("glmnet")
}
```
#### 5. Logistic model
```{r}

# Creating logistic regression model with 2 params tuned
make_logistic_model <- function() {
  logistic_reg(
    penalty = tune(),  # lambda
    mixture = tune()   # alpha
  ) |>
    set_engine("glmnet")
}
```
#### 6. LDA


```{r}
# LDA model
make_lda <- function() {
  discrim_linear() 
}
```

#### 7. QDA

```{r}
# QDA model
make_qda <- function() {
  discrim_quad() 
}
```

#### 8. Naive Bayes

```{r}
# Native Bayes model
make_naive_bayes <- function() {
  naive_Bayes() |>
  set_engine("naivebayes")
}
```

#### 9. Decision Tree - Classification


```{r}
# Classification decision tree model
make_class_tree <- function(minsplit = 5) {
  decision_tree() |>
    set_mode("classification") |>
    set_engine(
      "rpart",
      control = list(minsplit = minsplit)
    )
}

```

#### 10. Decision Tree - Regression - CCP for pruning


```{r}
# Regression decision tree model
make_reg_tree <- function() {
  # decision_tree() |>
  decision_tree(
    cost_complexity = tune()
  ) |>
    set_mode("regression") |>
    set_engine(
      "rpart"
    )
}

```

#### 11. Linear SVM model


```{r}
# linear SVM model
make_linear_svm_model <- function() {
  svm_linear(cost = tune()) |>
    set_mode("classification") |>
    set_engine("kernlab")
}
```

#### 12. Non-linear SVM model


```{r}
# Non-linear SVM model
make_rbf_svm_model <- function() {
  svm_rbf(
    cost = tune(),
    # rbf_sigma = tune()
    rbf_sigma = 1
  ) |>
    set_mode("classification") |>
    set_engine("kernlab")
}
```

#### 13. Neural net - Classification model


```{r}
# Neural net classification model
make_nn_class_model <- function() {
  mlp(
    hidden_units = tune()
  ) |>
    set_mode("classification") |>
    set_engine("nnet")
}
```

#### 14. Neural net - Regression model


```{r}
# Neural net regression model
make_nn_reg_model <- function() {
  mlp(
    hidden_units = tune()
  ) |>
    set_mode("regression") |>
    set_engine("nnet")
}
```
```{r}
# Neural net regression model with both penalty and hidden_units tuned
make_nn_reg_model_v2 <- function() {
  mlp(
    penalty = tune(), 
    hidden_units = tune()
  ) |>
    set_mode("regression") |>
    set_engine("nnet")
}
```
### Step 4 - Getting workflow


```{r}
# Creating workflow
make_workflow <- function(recipe, model) {
  workflow() |>
    add_model(model) |>
    add_recipe(recipe)
}
```

#### use


```{r}
#lm_wf <- make_workflow(
#  recipe = penguins_recipe,
#  model = make_lm_model()
#)
#
#lm_wf
```

### Step 4.1 - Getting validation sets

#### Step 4.1.1 - v fold cross validation


```{r}
# V-Fold cross validation
make_vfold_cv <- function(data,
                          v = 10,
                          repeats = 5,
                          strata = NULL) {

  strata <- enquo(strata)

  vfold_cv(
    data = data,
    v = v,
    repeats = repeats,
    strata = !!strata
  )
}
```

#### Step 4.1.2 - LOOCV cross validation


```{r}
# Leave-one-out-cross-validation
make_loocv <- function(data) {
  loo_cv(data)
}
```
#### Step 4.1.3 - Bootstrapping

```{r}
# Bootstrap cross validation
make_bootstraps <- function(data,
                            times = 100,
                            strata = NULL) {

  strata <- enquo(strata)

  bootstraps(
    data = data,
    times = times,
    strata = !!strata
  )
}
```
### Step 4.2 Getting tuned workflow options

#### Step 4.2.1 tuned ridge/lasso workflow

```{r}
# Function to get tuned results for ridge/lasso
get_ridgelasso_tuned <- function(workflow,
                                resamples,
                                penalty_min = -4,
                                penalty_max = 0,
                                levels = 20,
                                metric = "rmse") {
  penalty_grid <- grid_regular(
    penalty(range = c(penalty_min, penalty_max)),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = penalty_grid
  )

  tuned
}
```
```{r}
# Get finalized workflow for ridge/lasso. RMSE is used to choose 
# the best
get_ridgelasso_finalized_workflow <- function(workflow,
                                resamples,
                                penalty_min = -4,
                                penalty_max = 0,
                                levels = 20,
                                metric = "rmse") {
  penalty_grid <- grid_regular(
    penalty(range = c(penalty_min, penalty_max)),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = penalty_grid
  )

  best_params <- select_best(tuned, metric = metric)

  finalize_workflow(workflow, best_params)
}
```



* USE 

```{r}
#lasso_wf <-
#  make_workflow(
#    recipe = penguins_recipe,
#    model = make_lasso_model()
#  )
#
#final_lasso_wf <-
#  get_ridgelasso_finalized_workflow(
#    workflow = lasso_wf,
#    resamples = cv_splits
#  )
#
#lass_tuned <-
#  get_ridgelasso_tuned(
#    workflow = lasso_wf,
#    resamples = cv_splits
#  )
#autoplot(lass_tuned)
#show_best(lass_tuned, metric = "rmse", n = 20)
#
#lambda <- tibble(
# penalty = 0.01
#)
#
#final_lasso_wflow <- lasso_wf |>
# finalize_workflow(lambda)
#final_lasso_wflow
```



#### 4.2.2 Tuned elasticnet workflow

```{r}
# Get tuned results for elasticnet
get_elasticnet_tuned <- function(workflow,
                                resamples,
                                penalty_min = -4,
                                penalty_max = 0,
                                levels = 5,
                                metric = "rmse") {
  penalty_grid <- grid_regular(
    penalty(range = c(penalty_min, penalty_max)),
    mixture(),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = penalty_grid
  )

  tuned
}
```
```{r}
# Get finalised workflow for regression based on elasticnet
get_elasticnet_finalized_workflow <- function(workflow,
                                resamples,
                                penalty_min = -4,
                                penalty_max = 0,
                                levels = 5,
                                metric = "rmse") {
  penalty_grid <- grid_regular(
    penalty(range = c(penalty_min, penalty_max)),
    mixture(),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = penalty_grid
  )

  best_params <- select_best(tuned, metric = metric)

  finalize_workflow(workflow, best_params)
}
```



* USE 

```{r}
#elasticnet_wf <-
#  make_workflow(
#    recipe = penguins_recipe,
#    model = make_elasticnet_model()
#  )
#
#final_elasticnet_wf <-
#  get_elasticnet_finalized_workflow(
#    workflow = elasticnet_wf,
#    resamples = cv_splits
#  )
#
#elasticnet_tuned <-
#  get_elasticnet_tuned(
#    workflow = lasso_wf,
#    resamples = cv_splits
#  )
#autoplot(elasticnet_tuned)
#show_best(elasticnet_tuned, metric = "rmse", n = 20)
#
#lambda <- tibble(
#  penalty = 0.01,
#  mixture = 0.5
#)
#
#final_elasticnet_wflow <- elasticnet_wf |>
# finalize_workflow(lambda)
#final_elasticnet_wflow
```
#### 4.2.3 Tuned logistic regression workflow

```{r}
# Get finalized workflow for logistic regression. roc is used to select the best
get_logistic_finalized_workflow <- function(workflow,
                                resamples,
                                penalty_min = -4,
                                penalty_max = 0,
                                levels = 5,
                                metric = "roc_auc") {
  penalty_grid <- grid_regular(
    penalty(range = c(penalty_min, penalty_max)),
    mixture(),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = penalty_grid,
    metrics = metric_set(accuracy, roc_auc, pr_auc)
  )

  best_params <- select_best(tuned, metric = metric)

  finalize_workflow(workflow, best_params)
}
```

#### 4.2.4 Pruned Regression Workflow


```{r}

# Get finalized workflow for pruned regression decision tree. 
# rmse is used to select the best
get_pruned_tree_finalized_workflow <- function(workflow,
                                               resamples,
                                               cp_min = -5,
                                               cp_max = -1,
                                               levels = 10,
                                               metric = "rmse") {
  param_grid <- grid_regular(
    cost_complexity(range = c(cp_min, cp_max)),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = param_grid,
    metrics = metric_set(rmse)
  )

  best_params <- tuned |>
    select_by_one_std_err(
      metric = metric,
      desc(cost_complexity)
    )

  finalize_workflow(workflow, best_params)
}
```
#### 4.2.5 - Linear svm finalised workflow


```{r}
# Get finalized workflow for linear svm. 
# roc is used to select the best

get_linear_svm_finalized_workflow <- function(workflow,
                                              resamples,
                                              cost_min = -5,
                                              cost_max = 5,
                                              levels = 20,
                                              metric = "roc_auc") {
  cost_grid <- grid_regular(
    cost(range = c(cost_min, cost_max)),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = cost_grid,
    metrics = metric_set(accuracy, roc_auc, pr_auc)
  )

  best_params <- select_best(tuned, metric = metric)

  finalize_workflow(workflow, best_params)
}
```

#### 4.2.6 - Non-linear SVM finalized workflow


```{r}
# Get finalized workflow for non-linear svm. 
# roc is used to select the best

get_nonlinear_svm_finalized_workflow <- function(workflow,
                                                 resamples,
                                                 cost_min = -3,
                                                 cost_max = 3,
                                                 # sigma_min = -4,
                                                 # sigma_max = -1,
                                                 levels = 5,
                                                 metric = "roc_auc") {
  param_grid <- grid_regular(
    cost(range = c(cost_min, cost_max)),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = param_grid,
    metrics = metric_set(accuracy, roc_auc, pr_auc),
    control = control_grid(save_pred = TRUE)
  )

  best_params <- select_best(tuned, metric = metric)

  finalize_workflow(workflow, best_params)
}
```

#### 4.2.7 - NN Classification finalized workflow


```{r}
# Get finalized workflow for classification using neural net . 
# roc is used to select the best

get_nn_class_finalized_workflow <- function(workflow,
                                            resamples,
                                            hidden_units_min = 1,
                                            hidden_units_max = 20,
                                            levels = 5,
                                            metric = "roc_auc") {
  param_grid <- grid_regular(
    hidden_units(range = c(hidden_units_min,
                           hidden_units_max)),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = param_grid,
    metrics = metric_set(accuracy, roc_auc, pr_auc)
  )

  best_params <- select_best(tuned, metric = metric)

  finalize_workflow(workflow, best_params)
}
```

#### 4.2.8 - NN Regression finalized workflow


```{r}
# Get finalized workflow for regression using neural net . 
# rmse is used to select the best

get_nn_reg_finalized_workflow <- function(workflow,
                                          resamples,
                                          hidden_units_min = 1,
                                          hidden_units_max = 20,
                                          levels = 5,
                                          metric = "rmse") {
  param_grid <- grid_regular(
    hidden_units(
      range = c(hidden_units_min,
                hidden_units_max)
    ),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = param_grid,
    metrics = metric_set(rmse)
  )

  best_params <- select_best(tuned, metric = metric)

  finalize_workflow(workflow, best_params)
}
```

### 4.2.9 - NN Regression finalized workflow - V2 - tuned 2 vars

```{r}
# Get finalized workflow for regression using neural net . penalty and hidden units tuned.
# rmse is used to select the best

get_nn_reg_finalized_workflow_v2 <- function(workflow,
                                             resamples,
                                             hidden_units_min = 1,
                                             hidden_units_max = 20,
                                             penalty_range = c(-6, -1),
                                             levels = 5,
                                             metric = "rmse") {

  param_grid <- grid_regular(
    hidden_units(
      range = c(hidden_units_min,
                hidden_units_max)
    ),
    penalty(
      range = penalty_range
    ),
    levels = levels
  )

  tuned <- tune_grid(
    workflow,
    resamples = resamples,
    grid = param_grid,
    metrics = metric_set(rmse)
  )

  best_params <- select_best(
    tuned,
    metric = metric
  )

  finalize_workflow(
    workflow,
    best_params
  )
}
```
### Step 5 - Fitting Model


```{r}
# Getting model fit
fit_model <- function(workflow, training_data) {
  fit(workflow, data = training_data)
}
```
#### Use


```{r}
#fitted_wf <- fit_model(
#  workflow = lm_wf,
#  training_data = penguins_train
#)
#
#fitted_wf
```

### Step 6 - Predict


```{r}
# model augmented with test data
augment_model <- function(fit_object, new_data) {
  augment(fit_object, new_data = new_data)
}
```

#### Use


```{r}
#augmented_test <- augment_model(
#  fit_object = fitted_wf,
#  new_data = penguins_test
#)
#
#augmented_test
```

### Step 7 - Write Results in file


```{r}
# Write specified column into csv file
write_augmented_csv <- function(data,
                                 file_path,
                                 cols = NULL,
                                 col_names = NULL) {

  out <- data

  # Select columns if provided
  if (!is.null(cols)) {
    out <- out |> select(all_of(cols))
  }

  # Rename columns if provided
  if (!is.null(col_names)) {
    if (length(col_names) != ncol(out)) {
      stop("`col_names` must match number of selected columns.")
    }
    names(out) <- col_names
  }

  write_csv(out, file_path)

  invisible(out)
}
```

#### Use 


```{r}
#augmented_test |>
#  write_augmented_csv("predictions.csv")
#
#augmented_test |>
#  write_augmented_csv(
#    file_path = "predictions.csv",
#    cols = c("body_mass_g", ".pred")
#  )
#
#augmented_test |>
#  write_augmented_csv(
#    file_path = "predictions.csv",
#    cols = c("body_mass_g", ".pred"),
#    col_names = c("actual", "result")
#  )
#
#make_workflow(
#  model = make_lm_model(),
#  recipe = penguins_recipe
#) |>
#  fit_model(training_data = penguins_train) |>
#  augment_model(test_data = penguins_test) |>
#  write_augmented_csv(
#    file_path = "predictions.csv",
#    cols = c("body_mass_g", ".pred"),
#    col_names = c("actual", "result")
#  )
#

```


### Full Workflows

#### 1. lm model


```{r}
library(palmerpenguins)
data("penguins", package = "palmerpenguins")
head(penguins) |>
kable()

splits <-
  penguins |>
    make_data_split(train = 0.6)
head(splits$train)
out <-
  splits$train |>
    make_recipe(
      outcome = body_mass_g,
      id_vars = c("island", "sex", "year")
    ) |>
    make_workflow(
      model = make_lm_model()
    ) |>
    fit_model(
      training_data = splits$train
    ) |>
    augment_model(
      new_data = splits$test
    )

out |>
  write_augmented_csv(
    file_path = "test.csv",
    cols = ".pred",
    col_names = "result"
  )
  

```

#### 2. glm - elasticnet with tuning - Custom Recipe


```{r}
data(ames, package = "modeldata")
ames <- ames |>
 mutate(Sale_Price = log10(Sale_Price))

splits <-
  ames |>
    make_data_split(train = 0.8)

ames_recipe <- 
  recipe(Sale_Price ~ Gr_Liv_Area + Year_Built + Longitude + Latitude, data = splits$train) |>
  step_ns(Longitude, Latitude, deg_free = 5) |>
  step_normalize(all_numeric_predictors())

ames_folds <- 
  splits$train |>
    make_vfold_cv(repeats = 1)

out <-
    make_workflow(
      recipe = ames_recipe,
      model = make_elasticnet_model()
    ) |>
    get_elasticnet_finalized_workflow(
      resamples = ames_folds
    ) |>
    fit_model(
      training_data = splits$train
    ) |>
    augment_model(
      new_data = splits$test
    )

out |> rmse(Sale_Price, .pred)

```

#### 3. logistic_reg model - Not tuned

```{r}
data(german.credit, package = "fairml")
gcredit <- as_tibble(german.credit)

splits <-
  gcredit |>
    make_data_split(train = 0.8, strata = Credit_risk)

out <-
  splits$train |>
    make_recipe(
      outcome = Credit_risk,
      # id_vars = c("island", "sex", "year")
    ) |>
    make_workflow(
      model = logistic_reg() 
    ) |>
    fit_model(
      training_data = splits$train
    ) |>
    augment_model(
      new_data = splits$test
    )

glimpse(out)
```
###### Confusion matrix


```{r}
gcredit_cm <-
  out |>
    conf_mat(Credit_risk, .pred_class)
gcredit_cm
gcredit_cm |> summary()

```

#### 3. logistic_reg model - tuned

```{r}
data(german.credit, package = "fairml")
gcredit <- as_tibble(german.credit)

splits <-
  gcredit |>
    make_data_split(train = 0.8, strata = Credit_risk)

gcredit_folds <- 
  splits$train |>
    make_vfold_cv(strata = Credit_risk)

out <-
  splits$train |>
    make_recipe(
      outcome = Credit_risk,
    ) |>
    make_workflow(
      model = make_logistic_model() 
    ) |>
    get_logistic_finalized_workflow(
      resamples = gcredit_folds
    ) |>
    fit_model(
      training_data = splits$train
    ) |>
    augment_model(
      new_data = splits$test
    )

```
Testing metrics

```{r}

metric_set_en <- metric_set(accuracy, roc_auc, pr_auc)
out |> metric_set_en(Credit_risk, estimate = .pred_class, .pred_BAD )
levels(gcredit$Credit_risk)
```
#### 4. LDA/QDA


```{r}
data(german.credit, package = "fairml")
gcredit <- as_tibble(german.credit)

splits <-
  gcredit |>
    make_data_split(train = 0.8, strata = Credit_risk)

out <-
  splits$train |>
    make_recipe(
      outcome = Credit_risk,
      id_vars = c("Account_status", "Credit_history", "Purpose", "Savings_bonds", 
        "Present_employment_since", "Other_debtors_guarantors", "Property",
        "Other_installment_plans", "Housing", "Job", "Telephone", "Foreign_worker",
      "Gender")
    ) |>
    make_workflow(
      model = make_qda() 
    ) |>
    fit_model(
      training_data = splits$train
    ) |>
    augment_model(
      new_data = splits$test
    )

```
```{r}

metric_set_en <- metric_set(accuracy, roc_auc, pr_auc)
out |> metric_set_en(Credit_risk, estimate = .pred_class, .pred_BAD )
cm <- out |> conf_mat(truth = Credit_risk, estimate = .pred_class)
cm
cm |> summary()
```

#### 5. Decision Tree - Classification


```{r}
penguins <- penguins |> drop_na()

splits <-
  penguins |>
    make_data_split(train = 0.6, strata = species)

# gcredit_folds <- 
  # splits$train |>
    # make_vfold_cv(strata = Credit_risk)

out_fit <-
  splits$train |>
    make_tree_recipe(
      outcome = species,
      id_vars = c("island", "flipper_length_mm", "sex")
    ) |>
    make_workflow(
      model = make_class_tree() 
    ) |>
    fit_model(
      training_data = splits$train
    )
penguin_rpart <- out_fit |> extract_fit_engine()
penguin_rpart |> print()
penguin_rpart |> rpart.plot()

out <-
  splits$train |>
    make_tree_recipe(
      outcome = species,
      id_vars = c("island", "flipper_length_mm", "sex")
    ) |>
    make_workflow(
      model = make_class_tree() 
    ) |>
    # get_logistic_finalized_workflow(
      # resamples = gcredit_folds
    # ) |>
    fit_model(
      training_data = splits$train
    ) |>
    augment_model(
      new_data = splits$test
    )

out
metric_set_en <- metric_set(accuracy, roc_auc, pr_auc)
out |> metric_set_en(species, estimate = .pred_class, .pred_Adelie, .pred_Chinstrap, .pred_Gentoo)
cm <- out |> conf_mat(truth = species, estimate = .pred_class)
cm
cm |> summary()

```
#### Decision Tree - Pruned regression


```{r}
data("Hitters", package = "ISLR2")
hitters <-
 Hitters |> drop_na() |>
    mutate(Salary = log10(Salary)) |>
    tibble()

splits <-
  hitters |>
    make_data_split(train = 0.8)

hitters_folds <- 
  splits$train |>
    make_vfold_cv()


out <-
  splits$train |>
    make_tree_recipe(
      outcome = Salary
    ) |>
    make_workflow(
      model = make_reg_tree()
    ) |>
    get_pruned_tree_finalized_workflow(resamples = hitters_folds) |>
    fit_model(
      training_data = splits$train
    ) |>
    augment_model(
      new_data = splits$test
    )

out |> rmse(truth = Salary, estimate = .pred)
```

#### Linear SVM - Classification - tuned


```{r}
penguins <- penguins |>
 drop_na() |>
 mutate(
  y = case_when(
   species == "Gentoo" ~ "Gentoo (1)",
   .default = "Not Gentoo (0)",
   .ptype = factor(levels = c("Not Gentoo (0)", "Gentoo (1)"))
  ) 
 )

splits <-
  penguins |>
    make_data_split(train = 0.8, strata = y)

p_folds <- 
  splits$train |>
    make_vfold_cv(strata = y)

out <-
  splits$train |>
    make_recipe(
      outcome = y,
      id_vars = c("island", "flipper_length_mm", "sex", "species", "body_mass_g")
    ) |>
    make_workflow(
      model = make_linear_svm_model() 
    ) |>
    get_linear_svm_finalized_workflow(resamples = p_folds) |>
    fit_model(
      training_data = splits$train
    )|>
    augment_model(
      new_data = splits$test
    )

```

```{r}
out
metric_set_en <- metric_set(accuracy, roc_auc, pr_auc)
out |> metric_set_en(truth = y, estimate = .pred_class, `.pred_Not Gentoo (0)`)
cm <- out |> conf_mat(truth = y, estimate = .pred_class)
cm
cm |> summary()


```
#### Non Linear SVM - Classification - tuned

```{r}
penguins <- penguins |>
 drop_na() |>
 mutate(
  y = case_when(
   species == "Chinstrap" ~ "Chinstrap (1)",
   .default = "Not Chinstrap (0)",
   .ptype = factor(levels = c("Not Chinstrap (0)", "Chinstrap (1)"))
  ) 
)


splits <-
  penguins |>
    make_data_split(train = 0.8, strata = y)

p_folds <- 
  splits$train |>
    make_vfold_cv(strata = y)

out <-
  splits$train |>
    make_recipe(
      outcome = y,
      id_vars = c("island", "flipper_length_mm", "sex", "species", "body_mass_g")
    ) |>
    make_workflow(
      model = make_rbf_svm_model() 
    ) |>
    get_nonlinear_svm_finalized_workflow(resamples = p_folds) |>
    fit_model(
      training_data = splits$train
    )|>
    augment_model(
      new_data = splits$test
    )


out
metric_set_en <- metric_set(accuracy, roc_auc, pr_auc)
out |> metric_set_en(truth = y, estimate = .pred_class, `.pred_Not Chinstrap (0)`)
cm <- out |> conf_mat(truth = y, estimate = .pred_class)
cm
cm |> summary()

```

#### NN - Regression - tuned



```{r}
data(ames, package = "modeldata")

ames <- ames |>
 mutate(Sale_Price = log10(Sale_Price)) |> # deal with skewness in data by taking logs, this is now our response
 drop_na() # drop missing values


splits <-
  ames |>
    make_data_split(train = 0.8)

p_folds <- 
  splits$train |>
    make_vfold_cv()

out <-
  splits$train |>
    make_recipe_addpred(
      outcome = Sale_Price,
      # id_vars = c("island", "flipper_length_mm", "sex", "species", "body_mass_g")
      preds = c("Gr_Liv_Area", "Year_Built", "Longitude", "Latitude")
    ) |>
    make_workflow(
      model = make_nn_reg_model_v2() 
    ) |>
    get_nn_reg_finalized_workflow_v2(resamples = p_folds) |>
    fit_model(
      training_data = splits$train
    )|>
    augment_model(
      new_data = splits$test
    )

out
```
```{r}
metric_set_en <- metric_set(mse, rsq)
out |> metric_set_en(truth = Sale_Price, estimate = .pred)


```

