---
title: Advanced R
format: 
  html:
    css: styles.css
  pdf:
    include-in-header:
      text: |
        \input{preamble.tex}
---


```{r}
#| include: false
knitr::opts_chunk$set(
  results = "markup",
  collapse = TRUE,
  comment = ""
)
```
```{r}
## library to inspect objects
library(lobstr)
library(tibble)
```
### 1. Names and Values

- reassignment [does not create]{.warning} new objects.
    * Can use either `obj_addr` or `ref`

```{r}
x <- c(1,2,3)
y <- x
obj_addr(x)
obj_addr(y)
ref(x)
ref(y)
```


- strict about syntactic names: only [characters]{.key}, [underscore]{.key}, [digits]{.key} and [dots]{.key} are allowed.
Can [only]{.warning} start with [digits]{.warning} or [dot but not followed by a number]{.warning}. 
In order to use [non-syntactic]{.warning} identifiers, use [backticks.]{.warning}
    
```{r}
`_abc` <- 1
`_abc`
`10` <- 15
`10`
```
- use `make.names` to convert to valid values. 
    
- [copy-on-modify]{.key} objects are copied only when they are modified

    * Once duplicate object was created for y, it is the [single pointer]{.warning} to the new object. So, subsequent
changes [do not create]{.warning} any further objects. This is an example of [modify-in-place.]{.key}

```{r}
x <- c(1,2,3)
cat(tracemem(x), "\n")
y <- x # not here
y[[2]] <- 22
y[[3]] <- 24
untracemem(y)
```
- function arguments follow copy-on-modify rule on the passed parameter too.
    
    * if the passed parameter is returned back without modification, this would be the same pointer that was passed.

- vector objects are created with values in place whereas [lists]{.warning} have [pointers]{.warning} to 
values created separately.

```{r}
vec <- c(1,2,3)
l <- list(1,2,3)

ref(vec)
ref(l)
```
- [copy-on-modify for lists]{.key} When lists are modified, they are [shallow copied]{.warning}.
This means that the copied list points to the old objects that were not changed. Only change is to the object that was.

```{r}
l1 <- list(1,2,3)
l2 <- l1
l2[[2]] <- 22
ref(l1, l2)
```
- [copy-on-modify for data frames]{.key} Dataframes has one pointer for each column 
    
    * [if column is modified,]{.key} Vector corresponding to the changed column will be created. 
Then, shallow copy with pointer to the new column vector.
    * [if row modified,]{.key} All vectors copied. Then, shallow copy to the pointers of [each new]{.warning} column.

```{r}
df <- data.frame(a = 1:5, b = 11:15)
df1 <- df
df1[,2] <- df1[,2] * 2
ref(df, df1)
df2 <- df
df2[2,] <- df2[2,1] * 3
ref(df, df2)
```
- character vectors do not store values. They point to a [global pool of strings]{.warning}

```{r}
x <- c("a", "a", "abc", "ac")
ref(x, character = TRUE)
```

- [No use declaring an integer vector, since if any member is inadvertently allocated a double value,
it triggers a copy action whereby a double vector is created upcasting each value in the original 
]{.warning}
```{r}
x <- c(1L, 2L, 3L)
tracemem(x)
x[[3]] <- 4
untracemem(x)
```
- [Address extensions are not reported by `tracemem`]{.warning}. Note below that `x` was copied when 
we attempted to extend its size to 3.
```{r}
x <- list(1, 2)
tracemem(x)
x[[3]] <- 3
ref(x)
```
[Q. What happens when a list is allocated to itself as a member?]{.question}
shallow-copy of x pointing to the old copy of x. [Note: Ignore the extra copies internally created by
`ref` triggering the two extra `tracemem` messages.]{.warning}

```{r}
x <- list(1,2)
ref(x)
tracemem(x)
x[[2]] <- x
ref(x)
```
- Repeated objects are counted only once

```{r}
x <- runif(1e6)
obj_size(x)
obj_size(list(x,x,x))
```
- Cast `data.frame` to list using `as.list` [if it has to be updated in a list]{.warning}. 
[lists use internal C code]{.key} and updates the same data.frame object rather than creating one
for each iteration.
```{r}
x <- data.frame(matrix(runif(5 * 1e4), ncol = 5))
tracemem(x)
med <- vapply(x, median, numeric(1))

for (i in seq_along(med)) {
  x[[i]] <- x[[i]] + med[[i]]
}
untracemem(x)
```
  - key line here is the `as.list(x)`
```{r}
x <- data.frame(matrix(runif(5 * 1e4), ncol = 5))
tracemem(x)
x <- as.list(x)
med <- vapply(x, median, numeric(1))

for (i in seq_along(med)) {
  x[[i]] <- x[[i]] + med[[i]]
}
untracemem(x)
```
- [environments]{.key} are [always modified in place.]{.key}

```{r}
e1 <- rlang::env(a = 1, b = 2, c = 3)
e2 <- e1
e1$c <- 4
e2$c
```

### 2. Vectors

- Two types of vectors: [atomic vectors]{.key} and [lists]{.key}
- Types of atomic vectors: [logical]{.key}, [integer]{.key}, [double]{.key} and [character]{.key}. `integer` and `double` together are [numeric]{.key}. 
*raw* and *complex* are rare.
- `Inf`, `-Inf` and `NaN` are of type double. 
- `10` is double but `10L` is integer. 
- use [`typeof()`]{.key} and `length()` for vectors. [Do not use `mode()` or `storage.mode`]{.warning}
- use `is.na()` for check NA
    - every operation with NA leads to NA except for `| & ^`

```{r}
x <- c(5, NA, 10, NA)
x == NA
x | TRUE 
```
- Use `is.logical()`, `is.double()`, `is.integer()` and `is.character()` [but never `is.numeric()`, `is.vector()`, `is.atomic()`]{.warning} as 
they perform different functionality.
- coercion order : character > double > integer > logical
- failed coercion leads to `NA` 

#### 2.1 Attributes

::: {.definition}
Name-value pairs that deem special meanings to atomic vectors. `attr(x,"attname") <- "attval"` used to set individual ones, 
use `str(attributes(x))` to view, use `names(x) <- c("a","b","c")` and `dim(x) <- c(2,3)` to set the special `names` and `dim` attributes.
:::

[Q. How to check if a vector is atomic, matrix or array?]{.question} First check with `is.null(dim(x))`. If `true` then *vector*, else use `is.matrix()` or
`is.array()`

```{r}
x <- c(1:20)
is.null(dim(x))
is.matrix(x)

dim(x) <- c(4, 5)
is.null(dim(x))
is.matrix(x)

dim(x) <- c(2,5,2)
is.null(dim(x))
is.matrix(x)
is.array(x)

```
:::{.definition}
**S3 Atomic vectors** Vectors with `class` attribute. Ex. `factor`,`Date`, `POSIXct`, `difftime`. 
Some of these S3 vectors have other attributes as well.
:::

- `factor` are built over `integer` types, has `class` and `levels` attributes. [Non-existent
values]{.warning} can be part of the levels as well.

```{r}
x <- c("b", "a", "a", "a", "b")
table(x)
x_f <- factor(x, levels = c("a", "b", "c"))
table(x_f)
attributes(x_f)
```
- `Date`, `POSIXct` and `difftime` built over `double`. First two are no. of days and no. of seconds
since *1970-01-01* and the latter uses `unit` to store durations.
```{r}
today <- Sys.Date()
today
str(today)
dt <- as.Date("1970-02-01")
unclass(dt)

tm <- as.POSIXct("1970-01-02 12:00", tz = "UTC")
attributes(tm) <- NULL
tm

wk_1 <- as.difftime(20, units = "weeks")
str(wk_1)
```
- use `attributes(x) <- NULL` to get their undelying values.

[Q. How to convert a list to an atomic vector?]{.question} `unlist(lst)`. `unlist` strips **all** 
attributes and retains only the underlying value.

- use `as_tibble()` to convert from `data.frame` to `tibble`

#### `tibble` > `data.frame`

(@) tibble allows reference to columns defined during construction.
```{r}
df <- tibble(
x = 1:3,
  y = x * 3
)
df
attributes(df)
```
(@) single column selected `[` is [not converted automatically to vector]{.warning}.

```{r}
df <- data.frame(
x = 1:3,
y = 11:13
)
df[,"y"]
df <- as_tibble(df)
df[,"y"]
```
(@) Including list as a column requires less code and is printed better

```{r}
df <- tibble(
x = 1:3,
y = list(1:2, 1:3, 1:4)
)
df
```
[Q. How do we get a selected tibble column directly in the form of a vector]{.question} `df[["col1"]]`

```{r}
df <- data.frame(
x = 1:3,
y = 11:13
)
df[["y"]]
```

```{r}
a <- numeric()
```
- `as.matrix()` vs `data.matrix()` Matrices have the same atomic types. `as.matrix` converts into the largest type whereas `data.matrix` converts to `double`
